### 2018-03-17 paper/5-implementations.tex
Minor changes

We provide an implementation of \Make using our framework in
Fig.~\ref{fig-make-implementation}. \Make processes keys in a linear order based
on a topological sort (see~\S\ref{sec-concurrency} for parallel \Make). For
each key, it builds it if it is older than any of its dependencies. We capture
the persistent build information that \Make stores by a pair
\hs{(@@modTime,}~\hs{now)} comprising the \emph{file modification time} function
\hs{modTime}~\hs{::}~\hs{k}~\hs{->}~\hs{Time} and the \emph{current time} \hs{now}.
Setting aside the explicit manipulation of file modification times, which in
reality is taken care of by the file system, the function \hs{make} captures
the essence of \Make in a clear and precise manner.

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Persistent build information
type Time       = Integer
type MakeInfo k = (k -> Time, Time)
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Make build system
make :: Eq k => Build Applicative (MakeInfo k) k v
make = topological process
  where
    process key deps act = do
        (modTime, now) <- gets getInfo
        let dirty = or [ modTime dep > modTime key | dep <- deps ]
        when dirty $ do
            v <- act
            let newModTime k = if k == key then now else modTime k
            modify $ putInfo (newModTime, now + 1) . putValue key v
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Standard graph algorithms (implementation omitted)
transitiveClosure :: Eq a => (a -> [a]) ->  a  -> [a]
topologicalSort   :: Eq a => (a -> [a]) -> [a] -> [a]
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Topological dependency strategy
topological :: Eq k => (k -> [k] -> State (Store i k v) v -> State (Store i k v) ())
            -> Build Applicative i k v
topological process task key = execState $ forM_ chain $ \k -> do
    let fetch k = gets (getValue k)
    case task fetch k of
        Nothing  -> return ()
        Just act -> process k (deps k) act
  where
    deps  = dependencies task -- dependencies is defined in @\S\ref{secdeps}@
    chain = topSort deps (reachable deps key)
\end{minted}
\vspace{-2mm}
\caption{An implementation of \Make using our framework.}\label{fig-make-implementation}
\vspace{-2mm}
\end{figure}

@@

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Approximation of task dependencies
data DependencyApproximation k = SubsetOf [k] | Unknown
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Persistent build information
type CalcChain k = [k]
type ExcelInfo k = ((k -> Bool, k -> DependencyApproximation k), CalcChain k)
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Result of speculative task execution
data Result k v = MissingDependency k | Result v [k]
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Reordering dependency strategy (implementation omitted, 21 lines)
reordering :: Ord k
    => (k -> State (Store i k v) (Result k v) -> State (Store i k v) (Maybe (Result k v)))
    -> Build Monad (i, CalcChain k) k v
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Excel build system
excel :: Ord k => Build Monad (ExcelInfo k) k v
excel = reordering process
  where
    process key act = do
        (dirty, deps) <- gets getInfo
        let rebuild = dirty key || case deps key of SubsetOf ks -> any dirty ks
                                                    Unknown     -> True
        if not rebuild
            then return Nothing
            else do
                result <- act
                case result of
                    MissingDependency _ -> return ()
                    Result v _dynamicDependencies -> do
                        let newDirty k = if k == key then True else dirty k
                        modify $ putInfo (newDirty, deps) . putValue key v
                return (Just result)
\end{minted}
\vspace{-2mm}
\caption{An implementation of \Excel using our framework.}\label{fig-excel-implementation}
\vspace{-2mm}
\end{figure}

@@

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Determine whether a trace is relevant to the current state
traceMatch :: (Monad m, Eq k) =>
    (k -> Hash v -> m Bool) -> k -> [Trace k v] -> m [Hash v]
traceMatch check key ts = mapMaybeM f ts
    where f (Trace k dkv v) = do
                b <- return (key == k) &&^ allM (uncurry check) dkv
                return $ if b then Just v else Nothing
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Recursive dependency strategy
recursive :: Eq k =>
    (k -> (k -> State (Store i k v, [k]) v)
       -> State (Store i k v, [k]) (v, [k])
       -> State (Store i k v, [k]) ())
    -> Build Monad i k v
recursive process task key store = fst $ execState (ensure key) (store, [])
    where
        ensure key = do
            let fetch k = do ensure k; gets (getValue k . fst)
            done <- gets snd
            when (key `notElem` done) $ do
                modify $ \(s, done) -> (s, key:done)
                case track task fetch key of
                    Nothing -> return ()
                    Just act -> process key fetch act
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Shake build system
shake :: (Eq k, Hashable v) => Build Monad [Trace k v] k v
shake = recursive $ \key fetch act -> do
    traces <- gets (getInfo . fst)
    poss <- traceMatch (\k v -> (==) v . hash <$> fetch k) key traces
    current <- gets (getHash key . fst)
    when (current `notElem` poss) $ do
        (v, ds) <- act
        modify $ \(s, done) ->
            let t = Trace key [(d, getHash d s) | d <- ds] (getHash key s)
            in (putInfo (t : getInfo s) (putValue key v s), done)

\end{minted}
\vspace{-2mm}
\caption{An implementation of \Shake using our framework.}\label{fig-shake-implementation}
\vspace{-2mm}
\end{figure}

@@

\begin{description}
\item[\hs{traceMatch}] takes a list of all recorded \hs{Trace} values, the \hs{key} you are currently building, and a function \hs{check} which checks a specified dependency. It returns the \hs{result} field of all traces that match. Since \hs{check} is in an arbitrary monad, the function has to use \hs{allM}/\hs{&&^} instead of the more usual \hs{all}/\hs{&&} functions\footnote{These functions are all available in the \cmd{extra} library on Hackage.}.
\item[\hs{recursive}] defines the dependency ordering pattern. It requires a \hs{process} function that builds a key given a way to recursively build a dependency, and a way to run task to produce result. The main purpose of \hs{recursive} is to ensure that in a single run any key is not built twice -- for which it extends the \hs{State} monad with a list of \hs{done} keys.
\item[\hs{shake}] ties everything together. First it tests if the traces allow the current state of the target key. If not, it builds the key and updates the store. The only subtlty is that Shake calls \hs{fetch} on the keys while checking them -- building the last-recorded dependencies before checking them. One minor annoyance is that the \hs{State} has been extended and thus needs to be projected using \hs{fst} before it is used.
\end{description}

@@

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
bazel :: (Eq k, Hashable v) => Build Applicative (Traces k v) k v
bazel = topological $ \key ds act -> do
    s <- get
    let Traces traces contents = getInfo s
    poss <- traceMatch (\k v -> return $ getHash k s == v) key traces
    if null poss then do
        v <- act
        modify $ \s ->
            let t = Trace key [(d, getHash d s) | d <- ds] (getHash key s)
                ts = Traces (t : traces) (Map.insert (hash v) v contents)
            in putInfo ts (putValue key v s)
    else do
        when (getHash key s `notElem` poss) $
            modify $ putValue key (contents Map.! head poss)
\end{minted}
\vspace{-2mm}
\caption{An implementation of \Bazel using our framework.}\label{fig-bazel-implementation}
\vspace{-2mm}
\end{figure}

Now we have seen all three dependency schemes, we can directly reuse \hs{topological} to define \Bazel. Furthermore, as \Bazel is a tracing build system, we can reuse \hs{Trace} and \hs{traceMatch}, along with the \hs{Traces} type from \S\ref{sec-constructive-traces}. With these pieces in place, the implementation of \Bazel is given in Figure~\ref{fig-bazel-implementation}. We first figure out the possible results given the current state. If there are no recorded traces for the current dependencies we run it and record the results, otherwise grab a suitable result from the \hs{contents} cache.

The program presented above captures certain aspects of \Bazel, but the real implementation makes one important additional assumption on \hs{Task} -- namely that it is \textit{deterministic}. With that assumption \Bazel is able to create the result hash in a trace by hashing together the result hashes of the dependencies and the key -- as the mapping between dependencies and results is deterministic. As a consequence \Bazel can compute the results of traces locally, without looking at \hs{Traces} (potentially saving a roundtrip to the server). To model this change in the code would require storing an additional transient piece of information \hs{done} of type \hs{Map k (Hash v)}, storing the computed hashes so far this run. With that available, \hs{getHash key s} would become:

\begin{minted}{haskell}
hash (key, [ done Map.! d | d <- ds])
\end{minted}

And that new hash would have to be stored in \hs{done}.

@@

Using the abstractions and approaches built thus far, we have shown how to combine dependency scheme and change approach to reproduce existing build systems. In the attached materials we have implemented 9 build systems corresponding to all three dependency schemes, matched with all three change approaches. To us, the most interesting build system as yet unavailable would matching recursive ordering with constructive traces - providing a cloud-capable build system with minimality, cut-off and monadic dependencies. Using our framework it is possible to define and test such a system as per Figure~\ref{fig-cloudshake-implementation}.

@@

\subsection{Smarter \hs{[Trace]} Structures}\label{sec-smart-traces}

In the examples above, we have used \hs{[Trace k v]} to capture a list of traces -- however, using a list necessarily means that finding the right trace takes $O(n)$. For each of the \hs{Trace} based systems it is possible to devise a smarter representation, which we sketch below. Note that these implementations do not avoid calls to \hs{compute}, merely overheads in the build system itself.

\begin{enumerate}
\item Any system using verifying traces, e.g. \Shake, is unlikely to see significant benefit from storing more than one \hs{Trace} per key\footnote{There is a small chance of a benefit if the dependencies change but the result does not, and then the dependencies change back to what they were before.}. Therefore, such systems can store \hs{Map k (Trace k v)}, where the initial \hs{k} is the \hs{key} field of \hs{Trace}.
\item Any system using \hs{Applicative} dependencies can omit the dependency keys from the \hs{Trace} as they can be recovered from the \hs{key} field.
\item Any \hs{Applicative} system storing constructive traces, e.g. \Bazel, can index directly from the key and results to the output result - i.e. \hs{Map (k, [Hash v]) (Hash v)}. Importantly, assuming the traces are stored on a central server, the client can compute the key and the hashes of its dependencies, then make a single call to the server to retrieve the result hash. In this formulation we have removed the possibility for a single key/dependency state to map to multiple different hashes, e.g. on a non-deterministic build - something \Bazel already prohibits which is discussed more in \S\ref{sec-non-determinism}.
\item Finally, a \hs{Monad} system with constructive traces can be represented as \hs{Map k (Choice k v)}, assuming a definition of \hs{Choice} as:
\begin{minted}{haskell}
data Choice k v = Choice k (Map (Hash v) Choice)
                | Result (Hash v)
\end{minted}
Here the \hs{Choice} effectively encodes a tree, asking successive questions about keys, and taking different branches based on the resulting value, until it reaches a final result. Implementing this structure over client-server communication requires either a chatty interface with lots of round-trips per \hs{Choice} step, or sending over a part of the tree that is not subsequently explored.
\end{enumerate}

>>>>

We provide an implementation of \Make using our framework in
Fig.~\ref{fig-make-implementation}. \Make processes keys in a linear order based
on a topological sort (see~\S\ref{sec-parallelism} for parallel \Make). For
each key, it builds it if it is older than any of its dependencies. We capture
the persistent build information that \Make stores by a pair
\hs{(@@modTime,}~\hs{now)} comprising the \emph{file modification time} function
\hs{modTime}~\hs{::}~\hs{k}~\hs{->}~\hs{Time} and the \emph{current time} \hs{now}.
Setting aside the explicit manipulation of file modification times, which in
reality is taken care of by the file system, the function \hs{make} captures
the essence of \Make in a clear and precise manner.

% Type signature for process. Not sure we really want this, as it's quite long
% and is repeated in topological.
% process :: k -> [k] -> State@\,@(Store@\,@(MakeInfo@\,\blk{k}@)@\,\blk{k}@@\,\blk{v}@)@\,\blk{v}@ -> State@\,@(Store@\,@(MakeInfo@\,\blk{k}@)@\,\blk{k}@@\,\blk{v}@)@\,@(@@)
\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Persistent build information
type Time       = Integer
type MakeInfo k = (k -> Time, Time)
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Make build system
make :: Eq k => Build Applicative (MakeInfo k) k v
make = topological process
  where
    process key deps act = do
        (modTime, now) <- gets getInfo
        let dirty = or [ modTime dep > modTime key | dep <- deps ]
        when dirty $ do
            v <- act
            let newModTime k = if k == key then now else modTime k
            modify $ putInfo (newModTime, now + 1) . putValue key v
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Standard graph algorithms (implementation omitted)
reachable :: Eq a => (a -> [a]) ->  a  -> [a]
topSort   :: Eq a => (a -> [a]) -> [a] -> [a]
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Topological dependency strategy
topological :: Eq k => (k -> [k] -> State (Store i k v) v -> State (Store i k v) ())
            -> Build Applicative i k v
topological process task key = execState $ forM_ chain $ \k -> do
    let fetch k = gets (getValue k)
    case task fetch k of
        Nothing  -> return ()
        Just act -> process k (deps k) act
  where
    deps  = dependencies task -- dependencies is defined in @\S\ref{secdeps}@
    chain = topSort deps (reachable deps key)
\end{minted}
\vspace{-2mm}
\caption{An implementation of \Make using our framework.}\label{fig-make-implementation}
\vspace{-2mm}
\end{figure}

@@

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Approximation of task dependencies
data DependencyApproximation k = SubsetOf [k] | Unknown
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Persistent build information
type CalcChain k = [k]
type ExcelInfo k = ((k -> Bool, k -> DependencyApproximation k), CalcChain k)
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Result of speculative task execution
data Result k v = MissingDependency k | Result v [k]
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Reordering dependency strategy (implementation omitted, 21 lines)
reordering :: Ord k
    => (k -> State (Store i k v) (Result k v) -> State (Store i k v) (Maybe (Result k v)))
    -> Build Monad (i, CalcChain k) k v
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Excel build system
excel :: Ord k => Build Monad (ExcelInfo k) k v
excel = reordering process
  where
    process key act = do
        (dirty, deps) <- gets getInfo
        let rebuild = dirty key || case deps key of SubsetOf ks -> any dirty ks
                                                    Unknown     -> True
        if not rebuild
            then return Nothing
            else do
                result <- act
                case result of
                    MissingDependency _ -> return ()
                    Result v dynamicDependencies -> do
                        let newDirty k = if k == key then True else dirty k
                        modify $ putInfo (newDirty, deps) . putValue key v
                return (Just result)
\end{minted}
\vspace{-2mm}
\caption{An implementation of \Excel using our framework.}\label{fig-excel-implementation}
\vspace{-2mm}
\end{figure}

@@

@@

\begin{description}
\item[\hs{traceMatch}] takes a list of all recorded \hs{Trace} values, the \hs{key} you are currently building, and a function \hs{check} which checks a specified dependency. It returns the \hs{result} field of all traces that match. Since \hs{check} is in an arbitrary monad, the function has to use \hs{allM}/\hs{&&^} instead of the more usual \hs{all}/\hs{&&} functions\footnote{These functions are all available in the \cmd{extra} library on Hackage.}.
\item[\hs{recursive}] defines the dependency ordering pattern. It requires a \hs{process} function that builds a key given a way to recursively build a dependency, and a way to run task to produce result. The main purpose of \hs{recursive} is to ensure that in a single run no key is built twice~--~for which it extends the \hs{State} monad with a list of \hs{done} keys.
\item[\hs{shake}] ties everything together. First it tests if the traces allow the current state of the target key. If not, it builds the key and updates the store. The only subtlety is that Shake calls \hs{fetch} on the keys while checking them~--~building the last-recorded dependencies before checking them. One minor annoyance is that the \hs{State} has been extended and thus needs to be projected using \hs{fst} before it is used.
\end{description}

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Determine whether a trace is relevant to the current state
traceMatch :: (Monad m, Eq k)
    => (k -> Hash v -> m Bool) -> k -> [Trace k v] -> m [Hash v]
traceMatch check key ts = mapMaybeM f ts
    where f (Trace k dkv v) = do
                b <- return (key == k) &&^ allM (uncurry check) dkv
                return $ if b then Just v else Nothing
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Recursive dependency strategy
recursive :: Eq k => (k -> (k -> State (Store i k v, [k]) v)
                  -> State (Store i k v, [k]) (v, [k])
                  -> State (Store i k v, [k]) ())
                  -> Build Monad i k v
recursive process task key store = fst $ execState (ensure key) (store, [@@])
    where
        ensure key = do
            let fetch k = do ensure k; gets (getValue k . fst)
            done <- gets snd
            when (key `notElem` done) $ do
                modify $ \(s, done) -> (s, key:done)
                case track task fetch key of
                    Nothing -> return ()
                    Just act -> process key fetch act
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Shake build system
shake :: (Eq k, Hashable v) => Build Monad [Trace k v] k v
shake = recursive $ \key fetch act -> do
    traces <- gets (getInfo . fst)
    poss <- traceMatch (\k v -> (==) v . hash <$> fetch k) key traces
    current <- gets (getHash key . fst)
    when (current `notElem` poss) $ do
        (v, ds) <- act
        modify $ \(s, done) ->
            let t = Trace key [(d, getHash d s) | d <- ds] (getHash key s)
            in (putInfo (t : getInfo s) (putValue key v s), done)

\end{minted}
\vspace{-3mm}
\caption{An implementation of \Shake using our framework.}\label{fig-shake-implementation}
\vspace{1mm}
\end{figure}

@@

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
bazel :: (Eq k, Hashable v) => Build Applicative (Traces k v) k v
bazel = topological $ \key ds act -> do
    s <- get
    let Traces traces contents = getInfo s
    poss <- traceMatch (\k v -> return $ getHash k s == v) key traces
    if null poss then do
        v <- act
        modify $ \s ->
            let t = Trace key [(d, getHash d s) | d <- ds] (getHash key s)
                ts = Traces (t : traces) (Map.insert (hash v) v contents)
            in putInfo ts (putValue key v s)
    else do
        when (getHash key s `notElem` poss) $
            modify $ putValue key (contents Map.! head poss)
\end{minted}
\vspace{-2mm}
\caption{An implementation of \Bazel using our framework; \hs{topological} is
defined in Fig.~\ref{fig-make-implementation}.}\label{fig-bazel-implementation}
\vspace{-2mm}
\end{figure}

Now we have seen all three dependency schemes, we can directly reuse \hs{topological} to define \Bazel. Furthermore, as \Bazel is a tracing build system, we can reuse \hs{Trace} and \hs{traceMatch}, along with the \hs{Traces} type from \S\ref{sec-constructive-traces}. With these pieces in place, the implementation of \Bazel is given in Fig.~\ref{fig-bazel-implementation}. We first figure out the possible results given the current state. If there are no recorded traces for the current dependencies we run it and record the results, otherwise grab a suitable result from the \hs{contents} cache.

The program presented above captures certain aspects of \Bazel, but the real implementation makes one important additional assumption on \hs{Task} -- namely that it is \textit{deterministic}. With that assumption \Bazel is able to create the result hash in a trace by hashing together the result hashes of the dependencies and the key -- as the mapping between dependencies and results is deterministic. As a consequence \Bazel can compute the results of traces locally, without looking at \hs{Traces} (potentially saving a roundtrip to the server). To model this change in the code would require storing an additional transient piece of information \hs{done} of type \hs{Map}~\hs{k}~\hs{(Hash}~\hs{v)}, storing the computed hashes so far this run. With that available, \hs{getHash}~\hs{key}~\hs{s} would become:

\begin{minted}[xleftmargin=10pt]{haskell}
hash (key, [ done Map.! d | d <- ds])
\end{minted}

\noindent
And that new hash would have to be stored in \hs{done}.

@@

Using the abstractions and approaches built thus far, we have shown how to combine dependency scheme and change approach to reproduce existing build systems. In the attached materials we have implemented 9~build systems corresponding to all three dependency schemes, matched with all three change approaches. To us, the most interesting build system as yet unavailable would matching recursive ordering with constructive traces~--~providing a cloud-capable build system with minimality, cutoff and monadic dependencies. Using our framework it is possible to define and test such a system as per Fig.~\ref{fig-cloudshake-implementation}.

@@

\subsection{Smarter \hs{[Trace]} data structures}\label{sec-smart-traces}

In the examples above, we have used \hs{[Trace}~\hs{k}~\hs{v]} to capture a list of traces -- however, using a list necessarily means that finding the right trace takes $O(n)$. For each of the \hs{Trace} based systems it is possible to devise a smarter representation, which we sketch below. Note that these implementations do not avoid calls to \hs{compute}, merely overheads in the build system itself.

\begin{enumerate}
\item Any system using verifying traces, e.g. \Shake, is unlikely to see significant benefit from storing more than one \hs{Trace} per key\footnote{There is a small chance of a benefit if the dependencies change but the result does not, and then the dependencies change back to what they were before.}. Therefore, such systems can store \hs{Map}~\hs{k}~\hs{(Trace}~\hs{k}~\hs{v)}, where the initial \hs{k} is the \hs{key} field of \hs{Trace}.
\item Any system using \hs{Applicative} dependencies can omit the dependency keys from the \hs{Trace} as they can be recovered from the \hs{key} field.
\item Any \hs{Applicative} build system storing constructive traces, e.g. \Bazel, can index directly from the key and results to the output result~--~i.e. \hs{Map}~\hs{(@@k,}~\hs{[Hash}~\hs{v])}~\hs{(Hash}~\hs{v)}. Importantly, assuming the traces are stored on a central server, the client can compute the key and the hashes of its dependencies, then make a single call to the server to retrieve the result hash. In this formulation we have removed the possibility for a single key/dependency state to map to multiple different hashes, e.g. on a non-deterministic build~--~something \Bazel already prohibits which is discussed more in~\S\ref{sec-non-determinism}.
\item Finally, a \hs{Monad} build system with constructive traces can be represented as \hs{Map}~\hs{k}~\hs{(Choice}~\hs{k}~\hs{v)}, assuming a definition of \hs{Choice} as:
\begin{minted}[xleftmargin=10pt]{haskell}
data Choice k v = Choice k (Map (Hash v) Choice)
                | Result (Hash v)
\end{minted}
Here the \hs{Choice} encodes a tree, asking successive questions about keys, and taking different branches based on the answers, until it reaches a final result. Implementing this structure over client-server communication requires either a chatty interface with lots of round-trips per \hs{Choice} step, or sending over a part of the tree that is not subsequently explored.
\end{enumerate}

### 2018-03-17 paper/6-engineering.tex
Minor changes

% Q: {Isn't ``early aborting'' automatically handled by short-circuiting
% withing monads, i.e. \hs{Nothing} or \hs{Left e} cancel out all subsequent
% computation?}
% NM says: The build system could chose an early abort monad, but the tasks are free in the monad, so you can't get the encoding to cross between them.

\subsection{Concurrency}\label{sec-concurrency}

While we have given simple implementations assuming a single thread of execution,
all the build systems we address are actually concurrent. While it complicates
the model to add concurrency, the complications can be restricted exclusively
to the dependency strategy:

@@

\noindent
Handling such tasks is possible in our framework, but requires an adjustment of
the correctness definition (\S\ref{sec-build-correctness}): instead of requiring
that the result of recomputing the \hs{task} matches the value in the \hs{ideal}
store:
\[
\hs{Just}~\hs{(}\hs{getValue}~\hs{ideal}~\hs{k)}~\hs{==}~\hs{compute}~\hs{task}~\hs{ideal}~\hs{k}\]
\noindent
we now require that \hs{ideal} contains \emph{one possible result of recomputing}
the \hs{task}:
\[
\hs{Just}~\hs{(}\hs{getValue}~\hs{ideal}~\hs{k)}~\hs{`elem`}~\hs{compute}~\hs{task}~\hs{ideal}~\hs{k}
\]

>>>>

% Q: {Isn't ``early aborting'' automatically handled by short-circuiting
% withing monads, i.e. \hs{Nothing} or \hs{Left e} cancel out all subsequent
% computation?}
% NM says: The build system could chose an early abort monad, but the tasks are
% free in the monad, so you can't get the encoding to cross between them.

\subsection{Parallelism}\label{sec-parallelism}

While we have given simple implementations assuming a single thread of execution,
all the build systems we address can actually build independent keys in parallel.
While it complicates the model, the complications can be restricted exclusively
to the dependency strategy:

@@

\noindent
Handling such tasks is possible in our framework, but requires an adjustment of
the correctness definition (\S\ref{sec-build-correctness}): instead of requiring
that the result of recomputing the \hs{task} matches the value stored in the \hs{result}:
\[
\hs{Just}~\hs{(}\hs{getValue}~\hs{k}~\hs{result)}~\hs{==}~\hs{compute}~\hs{task}~\hs{result}~\hs{k}\]
\noindent
we now require that \hs{result} contains \emph{one possible result of recomputing}
the \hs{task}:
\[
\hs{Just}~\hs{(}\hs{getValue}~\hs{k}~\hs{result)}~\hs{`elem`}~\hs{compute}~\hs{task}~\hs{result}~\hs{k}
\]
